---
title: "How does binarizing weighted networks change network characteristics?"
author: "Joel Eliason"
date: "7/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
packages <- c("igraph", "R.matlab", "ggplot2", "reshape2")
lapply(packages, library, character.only = TRUE)
options(scipen = 999)
```

Here we'll be examining how different network metrics (as seen [here](./bct_table.pdf)) change as we binarize a weighted undirected network.
The network we will be examining here is a network of functional whole-brain coactivations, as measured with fMRI.

First, we parse the data from a .mat file into a dataframe. Then we'll take a look at different thresholds.

```{r}
df <- readMat("Coactivation_matrix.mat")

nnodes <- dim(df$Coactivation.matrix)[1]
nlinks <- sum(df$Coactivation.matrix != 0) / 2 # Here we divide by 2 since the matrix is symmetric
```
We can see that there are `r nnodes` nodes and `r nlinks` links in the weighted network.

Next, we'll convert it to an `igraph` object and start to look at thresholding.

```{r}
graph <- graph_from_adjacency_matrix(df$Coactivation.matrix, weighted = T, mode = "undirected", diag = F)
```

### Degree
We'll start by looking at how a very simple characteristic, the *degree distribution*, changes with threshold:

```{r, warning=FALSE, message=FALSE}
thresholds <- c(0.01,0.05,0.1)
binarize <- function(threshold, matrix) {
  bin.graph <- graph_from_adjacency_matrix(matrix > threshold, mode = "lower", diag = F)
  degrees <- degree(bin.graph)
  df <- data.frame(degrees)
  colnames(df) <- c(paste("Binarized at threshold = ", threshold))
  return(df)
}

bin.dfs <- lapply(thresholds, binarize, matrix=df$Coactivation.matrix)

degrees <- degree(graph)
df <- data.frame(degrees)
colnames(df) <- c("Original")



ggplot(melt(cbind(df, do.call(cbind, bin.dfs)), value.name = "Degree"), aes(Degree)) + 
  geom_histogram(aes(y=..density..), bins=30,
                 color="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") + 
  facet_wrap(.~variable, scales="free")
```

