---
title: "k-regular graphs"
author: "Joel Eliason"
date: "7/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
packages <- c("igraph", "network", "sna")
lapply(packages, library, character.only = TRUE)
```
This work is intended to explore the relationship between several rather simple
network characteristics and the size (number of vertices) of graphs that are
randomly generated as *k-regular* graphs, that is, graphs in which each vertex has
*k* neighbors. For example, as a first experiment, I will be examining if and how
the distribution of the measure of average *closeness centrality* of a network will 
change as I vary the size of the graph.

Of course, an important point is that this work is intentionally very exploratory,
so the opportunity to go down rabbit holes will be continually present.

Here we can see an example of a *k-regular* graph:

```{r}
g <- k.regular.game(20, 4)

l <- layout.kamada.kawai(g)
plot(g, layout=l)
```

As a first pass, let's take a look at a couple histograms of the average closeness
of the graph for a couple different network sizes, but with *k* fixed.
```{r, cache=TRUE}
N = 20
k = 4
m = 10000
set.seed(1)
hist.closeness <- function (N, m, k) {
  graphs <- replicate(m, igraph::k.regular.game(N, k), simplify = FALSE)


  close <- lapply(graphs, igraph::closeness, normalized = TRUE)
  close.mean <- unlist(lapply(close, mean))
  hist(close.mean, col="lightblue", xlim=c(min(close.mean),max(close.mean)),
       xlab=paste("Mean closeness, N = ", N), ylab="Frequency", main="", breaks = 30)
}

par(mfrow=c(1,3))
hist.closeness(N, m, k)

N = 50
hist.closeness(N, m, k)

N = 100
hist.closeness(N, m, k)

```

Let's examine some of these centralities using target plots:
```{r, warning=FALSE}
N = 20
graph <- igraph::k.regular.game(N, k)
A <- igraph::get.adjacency(graph, sparse=FALSE)
g <- network::as.network.matrix(A)
sna::gplot.target(g, closeness(g), main="Closeness",
                  circ.lab = FALSE,
circ.col="skyblue",
usearrows = FALSE,
vertex.col="blue",
edge.col="darkgray")
```


